\bibliographystyle{plain}
%\bibliography{reference}

\begin{thebibliography}{99}

% --- ASR & Transcription ---

\bibitem{adedeji2024sound}
Ayo Adedeji, Sarita Joshi, and Brendan Doohan.
\newblock The sound of healthcare: Improving medical transcription {ASR}
  accuracy with large language models.
\newblock {\em arXiv preprint arXiv:2402.07658}, 2024.

\bibitem{banerjee2024medasr}
Sourav Banerjee, Ayushi Agarwal, and Promila Ghosh.
\newblock High-precision medical speech recognition through synthetic data and
  semantic correction: {United-MedASR}.
\newblock {\em arXiv preprint}, December 2024.

\bibitem{efstathiadis2024diarization}
Georgios Efstathiadis, Vijay Yadav, and Anzar Abbas.
\newblock {LLM}-based speaker diarization correction: A generalizable approach.
\newblock 2024.

\bibitem{klusty2024clinical}
Mitchell~A. Klusty, W.~Vaiden Logan, Samuel~E. Armstrong, Aaron~D. Mullen,
  Caroline~N. Leach, Ken Calvert, Jeff Talbert, and V.~K.~Cody Bumgardner.
\newblock Toward automated clinical transcriptions.
\newblock 2024.

\bibitem{papadopoulos2022primock57}
Alex {Papadopoulos Korfiatis}, Francesco Moramarco, Radmila Sarac, and
  Aleksandar Savkov.
\newblock {PriMock57}: A dataset of primary care mock consultations.
\newblock In {\em Proceedings of the Workshop on Clinical Natural Language
  Processing}, 2022.

\bibitem{radford2023whisper}
Alec Radford, Jong~Wook Kim, Tao Xu, Greg Brockman, Christine McLeavey, and
  Ilya Sutskever.
\newblock Robust speech recognition via large-scale weak supervision.
\newblock In {\em Proceedings of the 40th International Conference on Machine
  Learning ({ICML})}, 2023.

\bibitem{tran2022asr}
Brian~D. Tran, Ramya Mangu, Ming Tai-Seale, Jennifer~Elston Lafata, and
  Kai Zheng.
\newblock Automatic speech recognition performance for digital scribes: A
  performance comparison between general-purpose and specialized models tuned
  for patient-clinician conversations.
\newblock 2022.

% --- FHIR Resource Synthesis ---

\bibitem{frei2025inferno}
Johann Frei, Nils Feldhus, Lisa Raithel, Roland Roller, Alexander Meyer, and
  Frank Kramer.
\newblock Inferno: End-to-end agent-based {FHIR} resource synthesis from
  free-form clinical notes.
\newblock {\em arXiv preprint}, 2025.

\bibitem{hong2017integrating}
Na~Hong, Andrew Wen, Feichen Shen, Sunghwan Sohn, Sijia Liu, Hongfang Liu,
  and Guoqian Jiang.
\newblock Integrating structured and unstructured {EHR} data using an
  {FHIR}-based type system: A case study with medication data.
\newblock In {\em AMIA Annual Symposium Proceedings}, 2017.

\bibitem{hong2020nlp2fhir}
Na~Hong, Andrew Wen, Feichen Shen, Sunghwan Sohn, Chen Wang, Hongfang Liu,
  and Guoqian Jiang.
\newblock Developing a scalable {FHIR}-based clinical data normalization
  pipeline for standardizing and integrating unstructured and structured
  electronic health record data.
\newblock {\em Journal of the American Medical Informatics Association},
  2020.

\bibitem{li2024fhirgpt}
Yikuan Li, Hanyin Wang, Halid~Z. Yerebakan, Yoshihisa Shinagawa, and Yuan Luo.
\newblock {FHIR-GPT} enhances health interoperability with large language
  models.
\newblock {\em NEJM AI}, 1(8), August 2024.

\bibitem{miller2023smart}
Timothy~A. Miller, Andrew~J. McMurry, James Jones, Daniel Gottlieb, and
  Kenneth~D. Mandl.
\newblock The {SMART} {Text2FHIR} pipeline.
\newblock 2023.

\bibitem{riquelme2025llm}
{\'A}lvaro {Riquelme Tornel}, Pedro {Costa del Amo}, and Catalina
  {Mart{\'i}nez Costa}.
\newblock Large language models for automating structured clinical data
  standardization: {HL7 FHIR} use case.
\newblock 2025.

% --- Clinical NLP & LLMs ---

\bibitem{brens2025semantic}
Rafael Brens, Yuqiao Meng, Luoxi Tang, and Zhaohan Xi.
\newblock Semantic {NLP} pipelines for interoperable patient digital twins from
  unstructured {EHR}s.
\newblock 2025.

\bibitem{builtjes2025leveraging}
Luc Builtjes, Joeran Bosma, Mathias Prokop, Bram {van Ginneken}, and
  Alessa Hering.
\newblock Leveraging open-source large language models for clinical information
  extraction in resource-constrained settings.
\newblock {\em Journal of the American Medical Informatics Association}, 2025.

\bibitem{saadat2025enhancing}
Saeed Saadat, Majid {Khalilizad Darounkolaei}, Mohsen Qorbani, Atefe Hemmat,
  and Sadaf Hariri.
\newblock Enhancing clinical documentation with {AI}: Reducing errors,
  improving interoperability, and supporting real-time note-taking.
\newblock {\em InfoScience Trends}, 2025.

% --- Standards ---

\bibitem{hl7fhir2022r4b}
{HL7 International}.
\newblock {HL7 FHIR} Release 4B (v4.3.0), 2022.
\newblock \url{https://hl7.org/fhir/R4B/}.

\end{thebibliography}
