\section{Results and Discussion}

\subsection{Evaluation Scope}

At the time of writing, the proposed pipeline had been executed end-to-end on all 15 PriMock57 consultations from Day~1, including \texttt{day1\_consultation15}, for which complete outputs were available across all six stages in \texttt{batch\_outputs/primock57\_pipeline}. The current results should therefore be interpreted as a batch evaluation of the fully processed Day~1 subset rather than as a benchmark over the complete 57-consultation corpus.

Unlike the earlier draft of this section, the present evaluation includes direct transcript and speaker-role benchmarking against PriMock57 reference annotations. Four result categories are now reported quantitatively: (1) end-to-end pipeline completion and FHIR validation, (2) transcript quality measured with word error rate (WER), (3) speaker-role accuracy derived from overlap between Step~4 segments and the doctor/patient TextGrid intervals, and (4) automatic characterization of the extracted clinical-resource outputs. Resource-level precision/recall and gold-standard FHIR fidelity are still out of scope because no manual extraction annotations have yet been created.

\subsection{End-to-End Pipeline Execution}

All 15 sampled consultations completed all six pipeline stages successfully, producing intermediate JSON outputs for ingestion, diarization, ASR, transcript post-processing, clinical extraction, and final validation. Across the batch, the total processed audio duration was 8894.16\,s (approximately 148.2\,min), with a mean consultation length of 592.94\,s. Step~1 reported a mean \textit{speech\_ratio} of 0.787, with values ranging from 0.717 to 0.846, indicating that the recordings were consistently speech-dense after pre-processing.

Step~2 identified two speakers in every processed consultation. When considering diarized segments of duration $\geq 0.1$\,s, the mean segment count was 260.3 per consultation, with a range from 196 to 425. The cumulative voiced segment duration across the batch was 6819.05\,s, which is broadly consistent with the speech-ratio estimates reported at ingestion time. These results show that the combined orchestration of audio pre-processing, diarization, CPU-based medical ASR, LLM refinement, LLM extraction, and programmatic FHIR construction is operationally feasible as an end-to-end batch pipeline.

\subsection{Audio and Speaker-Role Benchmarking}

Reference transcripts were constructed by merging the PriMock57 doctor and patient TextGrid annotations in timestamp order. Step~3 was evaluated as the raw ASR output, and Step~4 was evaluated separately as a cleaned-transcript ablation rather than as the primary ASR system. Speaker-role accuracy was computed only for non-empty Step~4 segments by assigning each predicted segment to the reference role with the largest temporal overlap.

\input{generated/audio_benchmark_summary_table.tex}

At corpus level, Step~3 achieved a WER of 0.3224, while Step~4 increased WER to 0.5946. The same pattern held at consultation level: mean WER rose from 0.3207 for Step~3 to 0.5845 for Step~4, and Step~4 did not improve WER in any of the 15 consultations. This result is important for interpreting the role of the LLM cleanup stage. Step~4 often improves readability and local coherence, but it does so by rewriting, compressing, and normalizing the transcript rather than by preserving lexical fidelity to the reference wording. Accordingly, it should be treated as a downstream semantic preprocessing step, not as an ASR enhancement stage.

Step~4 speaker-role assignment achieved an overall segment-level accuracy of 0.8963 across 2651 scoreable segments, with consultation-level accuracies ranging from 0.8367 to 0.9835. This is strong enough to show that role attribution is often recoverable from the current pipeline outputs, but it still leaves a non-trivial error rate in a clinically important distinction.

\input{generated/audio_benchmark_day_figure.tex}
\input{generated/audio_benchmark_consultation_wer_figure.tex}
\input{generated/audio_benchmark_role_figure.tex}

Figure~\ref{fig:audio-benchmark-day-summary} summarizes the currently available day-level benchmark. Because only Day~1 has been processed end-to-end so far, the plot contains a single day aggregate. Figure~\ref{fig:consultation-wer-comparison} shows that the gap between Step~3 and Step~4 WER is consistent across the full Day~1 batch. Figure~\ref{fig:consultation-role-accuracy} shows that role assignment is materially more stable than transcript-faithful cleanup, although several consultations still fall below 0.87 accuracy.

\subsection{FHIR Validation Outcomes}

The strongest structural result remains the validity of the generated FHIR bundles. In all 15 processed consultations, Step~6 returned \texttt{valid=true} with zero recorded validation issues. This indicates that all generated resources could be parsed successfully by the \texttt{fhir.resources} R4B models and that no schema-level errors were detected in required fields, value types, or bundle structure.

This result is non-trivial in the present setting. The pipeline begins from conversational audio, passes through multiple stochastic and error-prone stages, and ultimately constructs typed FHIR resources that must satisfy a strict interoperability schema. The absence of validation errors across the full processed batch therefore supports the architectural decision to constrain LLM output through Pydantic schemas and perform final resource assembly programmatically rather than asking the language model to emit raw FHIR JSON directly.

\begin{table}[ht]
\centering
\caption{Aggregate summary of end-to-end pipeline outputs for all 15 completed PriMock57 Day~1 consultations.}
\label{tab:results-summary}
\begin{tabular}{lc}
\hline
Metric & Value \\
\hline
Completed consultations & 15 \\
Validation pass rate & 15/15 (100\%) \\
Validation issues & 0 \\
Total duration (s) & 8894.16 \\
Mean duration (s) & 592.94 \\
Mean speech ratio & 0.787 \\
Speech-ratio range & 0.717--0.846 \\
Mean diarized segments ($\geq 0.1$\,s) & 260.3 \\
\texttt{Encounter} resources & 15 \\
\texttt{Condition} resources & 186 \\
\texttt{MedicationStatement} resources & 51 \\
\texttt{Observation} resources & 13 \\
\texttt{Procedure} resources & 6 \\
\texttt{Provenance} resources & 271 \\
\hline
\end{tabular}
\end{table}

\subsection{Automatic Extraction Characterization}

Because no manual semantic gold annotations currently exist, extraction outputs cannot yet be scored for correctness. A weaker but still informative analysis is therefore to characterize the composition, density, and traceability of the Step~5 outputs across the processed batch. This analysis does not estimate extraction precision or recall; instead, it quantifies how many resources are produced, how that mix varies between consultations, and whether the generated resources remain linked to transcript evidence.

\input{generated/extraction_summary_table.tex}

Across the 15 processed Day~1 consultations, Step~5 produced 256 clinical resources in total, excluding the mandatory \texttt{Encounter} resources: 186 \texttt{Condition}, 51 \texttt{MedicationStatement}, 13 \texttt{Observation}, and 6 \texttt{Procedure} resources. \texttt{Condition} dominated the output mix, accounting for 72.7\% of all extracted clinical resources, whereas observations and procedures remained relatively sparse. At consultation level, the total number of clinical resources ranged from 10 to 20, with a mean of 17.07 per consultation.

\input{generated/extraction_resource_count_figure.tex}

The composition of these bundles varied materially across consultations rather than collapsing to a fixed template. Some consultations yielded medication-heavy outputs, whereas others were almost entirely condition-focused. Observations were absent in 8 of the 15 consultations, and procedures were absent in 10, which is consistent with the conversational content of many primary-care encounters but also suggests that the extraction stage is much more willing to emit problem-oriented resources than management or measurement resources.

\input{generated/extraction_resource_density_figure.tex}

Normalizing by consultation duration gives the same qualitative picture. The mean output density was 1.79 clinical resources per minute, ranging from 0.75 to 2.23. This spread indicates that the extraction stage is not merely emitting a length-proportional number of resources. For example, consultations \texttt{01}, \texttt{03}, and \texttt{12} produced more than 2.0 resources per minute, whereas consultation \texttt{11} produced only 0.75 resources per minute. In other words, the extraction step appears responsive to local consultation content, but it also exhibits substantial variability in how aggressively it materializes that content into FHIR resources.

One strong result from this automatic analysis is traceability. All 256 clinical resources carried source-segment extensions linking them back to transcript evidence, which means that every extracted condition, medication, observation, and procedure remained auditable at segment level. This does not validate semantic correctness, but it does show that the current architecture preserves evidence linkage consistently across the entire processed batch.

At the same time, the batch results indicate a recurring tendency toward relatively granular \texttt{Condition} generation. Across several consultations, closely related symptoms, provisional diagnoses, and contextual clinical statements are represented as separate \texttt{Condition} resources. From a clinical-informatics perspective, such granularity may be useful for downstream auditability, but it also raises the possibility of over-segmentation, where closely related manifestations are represented as distinct problems rather than as attributes or contextual qualifiers of a smaller number of primary conditions. This issue should be addressed in future work through resource-level manual annotation and precision/recall analysis.

\subsection{Interpretation Relative to the Research Question}

The study asked whether an automated pipeline built from existing pretrained components could transform raw primary-care consultation audio into clinically meaningful, FHIR R4B-conformant resources without manual post-editing. The current results provide a qualified positive answer to the structural part of this question. Across all 15 Day~1 cases, the system consistently generated valid FHIR bundles containing multiple clinically relevant resource types and attached provenance links back to source transcript segments.

However, the benchmarked transcript results place an important limit on that claim. A Step~3 WER of 0.3224 is substantially worse than the 8.8--10.5\% range reported in controlled patient--clinician ASR evaluations from prior work, and the Step~4 cleanup stage further increases WER because it is optimized for semantic rewriting rather than literal transcript fidelity. The role metric is materially better at 0.8963, but even that leaves enough error to affect downstream attribution of symptoms, assessments, and management statements. The present evidence therefore supports the claim that end-to-end audio-to-FHIR transformation is technically feasible within the proposed architecture, and that the extraction stage produces structured, traceable outputs with non-trivial consultation-to-consultation variation, but it does not yet demonstrate that the resulting records are consistently accurate enough for production clinical use.

\subsection{Limitations of the Current Evaluation}

Several limitations constrain the interpretation of the present results. First, the completed end-to-end sample still covers only the 15 consultations from Day~1 currently processed in the batch output directory and therefore cannot support strong generalization across PriMock57 as a whole. Second, no gold-standard FHIR annotations are available for direct resource-level comparison, so the evaluation is presently strongest on structural validity and weakest on semantic fidelity. The automatic extraction analysis introduced here describes output composition and traceability, but it should not be mistaken for an accuracy benchmark. Third, the role metric is based on temporal overlap with doctor/patient TextGrid intervals rather than a dedicated word-level diarization benchmark. Fourth, terminology normalization is performed on a best-effort basis by the LLM without a dedicated ontology lookup service, which likely limits coding accuracy even when the free-text clinical meaning is broadly correct.

These limitations do not invalidate the current findings, but they define their scope. The contribution of the present work is therefore best understood as the demonstration of a working end-to-end research pipeline with initial transcript and role benchmarks, together with strong structural FHIR conformance on a public clinical-audio dataset.

\subsection{Implications for Future Evaluation}

The next stage of evaluation should extend from structural feasibility to semantic measurement. Four analyses would be especially valuable. First, the current transcript benchmark should be expanded from Day~1 to the remaining PriMock57 days so that the day-level plot becomes a genuine cross-day comparison rather than a single-group summary. Second, extraction outputs should be compared with a small manually annotated subset to estimate precision, recall, and error typology at the resource and attribute levels. Third, role evaluation should be refined toward a stricter word-level or utterance-level diarization benchmark. Fourth, terminology normalization and FHIR field accuracy should be measured directly on the same gold subset.

Such analyses would enable a more rigorous answer to the clinical usefulness question. At present, the evidence supports feasibility, auditability, and structural interoperability; future evaluation is needed to establish semantic reliability.
